---
title: "PML"
author: "Yan Feng"
date: "July 11, 2017"
output: 
  md_document:
    variant: markdown_github
---

## Practical Machine Learning Assignment

### Overview

The data collect information from accelerometers on the belt, forearm, arm, and
dumbell while doing 5 different ways. The goal is to predict the manner people
exercise throught the accelerometer readings.

### Analysis

- Load and clean data

The training and testing files are downloaded from the course website and saved
in the working directory. 

```{r load}
pml.train <- read.csv("pml-training.csv", na.strings = c("NA", ""))
pml.test <- read.csv("pml-testing.csv", na.strings = c("NA", ""))
str(pml.train)
```

The first 6 columns are user ID, name and timestamp, and are deleted

```{r remove}
pml.train <- pml.train[, c(-6:-1)]
pml.test <- pml.test[, c(-6:-1)]
```

Since many columns contain NA that has no information, only columns with all
non-NA data are kept.

```{r keep}
pml.train <- pml.train[, apply(!is.na(pml.train), 2, all)]
pml.test <- pml.test[, apply(!is.na(pml.test), 2, all)]
```

- Modeling

This is a classification problem. 

Random forest is used to model the data. Cross validation of 5 fold is used to 
tune the model

Initially, all sample are used to fit the model, but it takes very long time.

The training data is split into 2 sets: one for the train (10%), and the other
for the testing (90%).

```{r split}
library(caret)
InTrain <- createDataPartition(y = pml.train$classe, p = 0.1, list = FALSE)
pml.train.1 <- pml.train[InTrain, ]
pml.train.2 <- pml.train[-InTrain,]
```

Fit random forest model with cross validation of 5 fold
```{r fit, cache=TRUE}
library(randomForest)
model.rf <- train(classe ~., data = pml.train.1, method = "rf", trControl = trainControl(method = "cv", number = 5), prox = TRUE, allowParallel = TRUE)
print(model.rf)
```

The final model uses mtry of 27. The accuray is 95.4%.

Use this model for the testing data.
```{r test}
pred <- predict(model.rf, pml.train.2)
sum(pred == factor(pml.train.2$classe)) / length(pred)
```

The testing accuracy is 96.0%

- Predict new data

```{r predict}
pred.new <- predict(model.rf, pml.test)
pred.new
```

- plot of 2 featurs, roll_dumbbell and pitch_dumbbell, of the new data 

```{r plot}
qplot(roll_dumbbell, pitch_dumbbell, col = pred.new, data = pml.test) + geom_point()
```

### Conclusion

1. The data is fit into a random forest model
2. The accuracy is 95.4% for the training set and 96.0% for the testing set
3. New data set is predicted
